---
title: "GIS and Spatial Data Management for a Pskem River Watershed"
author: "Timur Sabitov"
organization: "U.S. Geological Survey"
output:
  pdf_document: default
  html_document: default
---

```{r }
```
First, clean work space and load required libraries

```{r, message = FALSE}
rm(list=ls(all=T)) 
require (sp)
require(raster)
require (rgdal)
require (maptools)
require (OasisR)
```

Next, load a watershed polygon, as well as polygons of glaciers.
Following code reads shape files of polygons.
```{r, message = FALSE, include=FALSE}
#Important
# You can use option file.choose, to navigate into directory and load file this way.
# You can perform similar operation for all other input files in this function.
# myFile <- file.choose()
# watershedPoly <- readOGR(myFile)

#Personal upload of shape files.
watershedPoly <- readOGR("D:/WorkingDir/Mapping_Projects/Thesis_pskem/DA_Pskem.shp")
glaciersPoly <- readOGR("D:/WorkingDir/Mapping_Projects/Thesis_pskem/Landcover.shp")
```

Let's take look at plot of watershed boundary overlayed with glaciers.
```{r}
plot (watershedPoly)
plot (glaciersPoly, add = T, col = "lightblue")
```

Ok, let's load rest of the files for spatial analysis.
```{r, message = FALSE}
#Load DEM
DEM <- raster ("D:/WorkingDir/Mapping_Projects/Thesis_pskem/DEM/PSKEM_DEM_Clipped_UTM1.tif")
#Load band 10 & 11 into R, from Landsat 8
band_10 <- raster("D:/WorkingDir/Mapping_Projects/Thesis_pskem/Landsat_B10_B11/band10.tif") #change image name accordingly
band_11 <- raster("D:/WorkingDir/Mapping_Projects/Thesis_pskem/Landsat_B10_B11/band11.tif") #change image name accordingly
```

Check projections and resolution.
```{r}
watershedPoly
glaciersPoly
DEM
band_10
band_11
```

Ok, projections are different. We need to reproject shape files.
This is how it is done.
```{r}
#Reproject shapefiles into the same projection as DEM raster file
watershedPoly <- spTransform(watershedPoly, crs(DEM))
glaciersPoly <- spTransform (glaciersPoly, crs (DEM))

#Raster projection works same way but using projectRaster function
#projected_raster <- projectRaster(r, crs = sr)# Project Raster
```

Next step is to clip raster by polygon shape
```{r}
#Set extent of the raster to clip and clip it to watershed.
myExtent <- extent (watershedPoly)
DEM <- crop (DEM, myExtent)
band_10 <- crop (band_10, myExtent)
band_11 <- crop (band_11, myExtent)

#To remove pixels outside of the polygon, mask raster with polygon.
DEM <- mask (DEM, watershedPoly)
band_10 <- mask (band_10, watershedPoly)
band_11 <- mask (band_11, watershedPoly)
```

Check files one more time.
```{r}
band_10
band_11
DEM
```

As you can see, cell values are slightly off, we need to correct them as well.
```{r}
#Extents are different, so as resolution, we need to resample data.
band_10 = resample(band_10, DEM, "bilinear") #Just to be quicker, we resample to 27, instead one should consider resampling to 30m.
band_11 = resample(band_11, DEM, "bilinear")
```

Radiance to temperature conversion.
```{r}
#Convert radiance into temperature values
RADIANCE_MULT_BAND_10 <- 3.3420E-04 #Values from Metafile
RADIANCE_MULT_BAND_11 <- 3.3420E-04

RADIANCE_ADD_BAND_10 <- 0.10000
RADIANCE_ADD_BAND_11 <- 0.10000

#Calculate TOA from DN:
toa_band10 <- calc(band_10, fun=function(x){RADIANCE_MULT_BAND_10 * x + RADIANCE_ADD_BAND_10})
toa_band11 <- calc(band_11, fun=function(x){RADIANCE_MULT_BAND_11 * x + RADIANCE_ADD_BAND_11})

#Values from Metafile
K1_CONSTANT_BAND_10 <- 774.8853
K1_CONSTANT_BAND_11 <- 480.8883
K2_CONSTANT_BAND_10 <- 1321.0789
K2_CONSTANT_BAND_11 <- 1201.1442

#Calculate LST in Kelvin for Band 10 and Band 11
temp10_kelvin <- calc(toa_band10, fun=function(x){K2_CONSTANT_BAND_10/log(K1_CONSTANT_BAND_10/x + 1)})
temp11_kelvin <- calc(toa_band11, fun=function(x){K2_CONSTANT_BAND_11/log(K1_CONSTANT_BAND_11/x + 1)})

#Convert Kelvin to Celsius for Band 10 and 11
temp10_celsius <- calc(temp10_kelvin, fun=function(x){x - 273.15})
temp11_celsius <- calc(temp11_kelvin, fun=function(x){x - 273.15})
```

To extract slope and aspect from DEM.
```{r}
x <- terrain(DEM, opt=c('slope', 'aspect'), unit='degrees')
```

To create a stack of rasters.
```{r}
s <- stack(temp10_celsius, temp11_celsius, x$slope, x$aspect, DEM)
#Plot stacks.
plot (s)
```

To create a user defined function that applied to each cell of the raster inside of the polygons.
```{r}
multi.fun <- function(x, na.rm = T) {
  c(min = min(x, na.rm = T), mean = mean(x, na.rm = T), max = max(x, na.rm = T), sd = sd(x, na.rm =  T))
}
```

To extract values of function for each polygon of the glacier.
```{r}
Output <- extract (s, glaciersPoly, fun = multi.fun, df = TRUE, na.rm=FALSE)
head (Output)
```

If you need area and perimeter of the polygon.
```{r}
area (glaciersPoly)
perimeter (glaciersPoly)
```

Well done!